---
title: "Practical Machine Learning. Find a model to predict exercise maner"
output: html_document
---
##Summary

The goal of the analysis was to find a model to predict exercise maner based on data 
from project "Human Activity Recognition" ( http://groupware.les.inf.puc-rio.br/har ).
The initial dataset consist of 19622 observation of 160 variables. Before we built
a model, raw data was processed - variables with large amounts NA and empty values 
was removed from dataset. Some another variables was also removed.  
After data processing new dataset had only 55 variables.
We found multiclass prediction problem and Random Forest method was choosen. 
This method uses bootstrap for cross validation and could give us model with high accuracy rate.  
Before to train model we splitted the dataset to train and test samples what help us 
to check accuracy of the final model.  
We need need accuracy more than 97.5% (or 19.5 correct answers for 20 samples).
After model was trained, we found the model with training accuracy rate 99.4%.
We checked accuracy rate with our test sample.
For test semple the model had similar 99.7% accuracy, that is very good and was 
enough for correct prediction of assignment's test.
Thus we could conclude that we found the model, that could be used as a predictor for 
exercise maner with high accuracy.  
 
## Data processing
Before make any analysis we need to make some changes in the dataset.   
Initial data had many variables with empty values or NA values.
We desided to remove those variables from dataset. 
```{r, echo = TRUE, cache = TRUE}
rawdata <- read.csv("pml-training.csv")
countNA <- sapply(rawdata, function(x) sum(length(which(is.na(x)|x==""))))  
badcolumns <- countNA > 19000
prData <- rawdata[ ,!badcolumns]
```
Also we decided that first 5 variables (row number, user_name and 3 timestamps variables) 
could not be prediction factors. These variables don't characterize the certain activity, 
just the moment when data was collected 

```{r, echo = TRUE, cache = TRUE}
prData <- prData[ , c(-1:-5)]
```


## Build a model
We splitted our data for two datasets train and test. First dataset we used for model 
training and second for check model accuracy. 
 
```{r, echo = TRUE, cache = TRUE}
library(caret)
set.seed(3412)
trList <- createDataPartition(y=prData$classe, p=.6, list = FALSE)
trData <- prData[trList,]
teData <- prData[-trList,]
```

Next we should find multiclass prediction model. Random Forest method was choosen. 
This method allow to make multiclass prediction and uses bootstrap for cross validation 
and could give us model with high accuracy rate. Before to prepare this repord we builded 
few different models. To make process of final model training faster we decided to 
change training parameters. Parametr "mtry" was set to the best value 27 for our dataset.    

```{r, echo = TRUE, cache = TRUE}
newGrid = expand.grid(mtry = c(27))
trModel <- train(classe~., data = trData,method="rf",prox=TRUE, tuneGrid = newGrid) 
```

We get model with good accurancy

```{r, echo = TRUE, cache = TRUE}
trModel
```

## Check the model
We cheked accuracy of the model with our test dataset, that was prepared.
```{r, echo = TRUE, cache = TRUE}
testpredict <- predict(trModel,teData)
predright <- testpredict==teData$classe
table(testpredict,teData$classe)
```

To complete assignment we should to predict factor "classe" for 20 samples. 
We expected that model's accurancy rate should be more than 97.5% (or 19.5 correct answers for 20 samples)
Our model let us good error rate **`r round((sum(!predright)/dim(teData)[1])*100,1)` %** and accurancy rate **`r round((1-sum(!predright)/dim(teData)[1])*100,1)` %**   

We used the model to complete assignment 

```{r, echo = TRUE, cache = TRUE}
testdataset <- read.csv("pml-testing.csv")
prediction <- predict(trModel,testdataset)
```

```{r, echo = FALSE, cache = FALSE}
pml_write_files = function(x){
        n = length(x)
        for(i in 1:n){
                filename = paste0("problem_id_",i,".txt")
                write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
        }
}

#pml_write_files(prediction)
#write.table(prediction,file="fullresult.txt",quote=FALSE,row.names=FALSE,col.names=FALSE)
```
